### 2013-07-24 doc/asplos14/conclude.tex

Irregular computations are both important and challenging to execute
quickly.  Scaling these applications easily on commodity hardware has
been a historical challenge. \Grappa is a runtime framework that
simplifies this task for software developers and compiler
writers. This paper describes the \Grappa framework and its three main
components: a task library, a distributed shared memory system, and a
network aggregator. \Grappa's key aspect is extreme latency tolerance,
which not only hides network latency but also enables the system to
spend time on sophisticated work stealing and network optimizations,
trading latency for even more throughput.

Our evaluation of \Grappa reveals that the core components, scheduling
and communication, achieve their design goals.  Thousands of workers
can be efficiently context switched on a multicore processor, up to
the DRAM bandwidth.  Aggregating messages enables \Grappa to achieve
over 1.0~GUPS on 64~nodes.  We also explored four other algorithms:
unbalanced tree search, breadth first search, PageRank, and integer
sort, comparing performance of these algorithms on \Grappa, the Cray
XMT, and optimized MPI implementations.

Depending on the benchmark, \Grappa is between 11X faster and 5.4X
slower than MPI and UPC, and between 2.6X faster and 4.3X slower than
the XMT. When \Grappa is slower than MPI, we find that the MPI
implementation includes a specialized implementation of a feature
\Grappa provides generally to all applications. This generality comes
at a performance cost but eases the application developer's
task. Moreover, there are significant tuning opportunities the \Grappa
runtime system.  When cost-performance is considered, \Grappa on mass-market
x86-based clusters is a highly attractive option.

>>>>

Irregular computations are both important and challenging to execute
quickly.  Scaling these applications easily on commodity hardware has
been a historical challenge. \Grappa 
simplifies this task for software developers and compiler
writers. 
% This paper describes \Grappa's framework and its main
% components: a task library, a distributed shared memory system, and a
% network aggregator. 
\Grappa's key aspect is extreme latency tolerance,
which not only hides network latency but also enables the system to
spend time on sophisticated work stealing and network optimizations,
trading latency for even more throughput.

Our evaluation of \Grappa shows that the core components, scheduling
and communication, achieve their design goals.  Thousands of workers
can be efficiently context switched on a multicore processor, limited by DRAM bandwidth.  Aggregating messages enables \Grappa to achieve
over 1.0~GUPS on 64~nodes.  We also explored four other algorithms:
unbalanced tree search, breadth first search, PageRank, and integer
sort, comparing performance of these algorithms on \Grappa, the Cray
XMT, and optimized MPI implementations. \Grappa is between 11X faster and 5.4X
slower than MPI and UPC, and between 2.6X faster and 4.3X slower than
the XMT. When \Grappa is slower than MPI, we find that the MPI
implementation includes a specialized implementation of a feature
\Grappa provides generally to all applications. This generality comes
at a performance cost but eases the application developer's
task. Moreover, there are significant tuning opportunities the \Grappa
runtime system.  When cost-performance is considered, \Grappa on mass-market
x86-based clusters is a highly attractive option.

### 2013-07-24 doc/asplos14/relatedwork.tex

\Grappa's closest ancestor is the Threaded Abstract
Machine~\cite{CullerGSvE93}. This was a software runtime system
designed as a prototyping platform for dataflow execution models on
distributed memory supercomputers, and contained support for
inter-node communication and management of the memory hierarchy as
well as context switching and scheduling computation. The Active
Messages~\cite{vonEicken:1992:AMM:139669.140382} work that grew out of
this project inspired our communication layer. One of the conclusions
of this work~\cite{Culler:1993:TFL:647025.714362} was that context
switch costs can be low only when contexts are in cache, and that
latency tolerance was not sufficient to guarantee performance on
commodity processors. \Grappa demonstrates that times have changed:
modern commodity processors have sufficient bandwidth and prefetch
capacity to stream contexts from DRAM and sustain a very large number of active contexts.

@@

\vspace{0.5ex}
\noindent{\bf Distributed graph processing systems.} 
%\section{Distributed graph processing systems} 
While \Grappa is a general runtime system for any large-scale
concurrent application, it has been designed to perform especially
well on graph analysis. Other distributed graph processing frameworks
include Pregel~\cite{pregel:2010} and Distributed
GraphLab~\cite{distgraphlab:vldb12}. Pregel adopts a bulk-synchronous
parallel (BSP) execution model, which makes it inefficient on
workloads that could prioritize vertices. GraphLab overcomes this
limitation with an execution model that schedules vertex computations
individually, allowing prioritization, which gives faster convergence
in a variety of iterative algorithms. GraphLab, however, imposes a
rigid model that requires programmers to express computation as
transformations on a vertex and its edge list, with information only
from adjacent vertexes. Pregel is only slightly less restrictive, as
the input data can be any vertex in the graph. \Grappa also supports
dynamic parallelism with asynchronous execution, but parallelism is
expressed as tasks or loop iterations, which is a far more general
programming model for irregular computation tasks.
PowerGraph~\cite{powergraph:osdi12} improves the performance of
GraphLab for real-world graphs with power-law degree distributions by
using a vertex cut for graph partitioning. Algorithmic transformations
like this  and would also improve the
performance of graph applications on \Grappa.

>>>>

\Grappa's closest ancestor is the Threaded Abstract
Machine~\cite{CullerGSvE93}. This was a software runtime system
designed as a prototyping platform for dataflow execution models on
distributed memory supercomputers, and contained support for
inter-node communication and management of the memory hierarchy as
well as context switching and scheduling computation. The Active
Messages~\cite{vonEicken92} work that grew out of
this project inspired our communication layer. One of the conclusions
of this work~\cite{Culler:1993:TFL:647025.714362} was that context
switch costs can be low only when contexts are in cache, and that
latency tolerance was not sufficient to guarantee performance on
commodity processors. \Grappa demonstrates that times have changed:
modern commodity processors have sufficient bandwidth and prefetch
capacity to stream contexts from DRAM and sustain a very large number of active contexts.

@@

\vspace{0.5ex}
\noindent{\bf Distributed graph processing systems.} 
%\section{Distributed graph processing systems} 
While \Grappa is a general runtime system for any large-scale
concurrent application, it performs especially
well on graph analysis. Other distributed graph processing frameworks
include Pregel~\cite{pregel:2010} and Distributed
GraphLab~\cite{distgraphlab:vldb12}. Pregel adopts a bulk-synchronous
parallel (BSP) execution model, which makes it inefficient on
workloads that could prioritize vertices. GraphLab overcomes this
limitation with an execution model that schedules vertex computations
individually, allowing prioritization, which gives faster convergence
in a variety of iterative algorithms. GraphLab, however, imposes a
rigid model that requires programmers to express computation as
transformations on a vertex and its edge list, with information only
from adjacent vertexes. Pregel is only slightly less restrictive, as
the input data can be any vertex in the graph. \Grappa also supports
dynamic parallelism with asynchronous execution, but parallelism is
expressed as tasks or loop iterations, which is a far more general
programming model for irregular computation tasks.
PowerGraph~\cite{powergraph:osdi12} improves the performance of
GraphLab for real-world graphs with power-law degree distributions by
using a vertex cut for graph partitioning. Algorithmic transformations
like this  and would also improve the
performance of graph applications on \Grappa.
